Basic concepts and tools for the Toki Pona minimalist and constructed language: Wordnet synsets; analysis of the vocabulary; synthesis and syntax highlighting of texts

### 2017-12-16 article/article.tex
detail about 24-bit true color

\begin{document}
\title{Basic concepts and tools for the Toki Pona minimalist and
constructed language:
Wordnet synsets; analysis of the vocabulary; synthesis and syntax highlighting of texts}
% Anthropological computation
\author{Renato Fabbri\\
\texttt{renato.fabbri@gmail.com}\\
University of São Paulo,\\
Institute of Mathematical and Computer Sciences\\
São Carlos, SP, Brazil
}
\maketitle
\begin{abstract}
A minimalist constructed language (conlang)
is useful for experiments and comfortable for making tools.
The Toki Pona (TP) conlang is minimalist both in the vocabulary
(with only 14 letters and 124 words)
and in the $\approx10$ syntax rules.
The language is useful for being a used and somewhat established
minimalist conlang with at least hundreds of fluent speakers.
In this article, we describe current concepts and resources
for TP,
and make available Python scripted routines for
the analysis of the language,
the synthesis of texts, the specification of syntax highlighting
schemes, and the achievement of a preliminary TP
Wordnet~\cite{wordnet}.
We focus on the analysis of the basic vocabulary,
as corpus analyses were found in~\cite{corpus}.
The synthesis is based on sentence templates,
relates to context by keeping track of used words,
and renders larger texts by using a fixed number of phonemes (e.g. for poems)
and number of sentences, words and letters (e.g. for paragraphs).
Syntax highlighting 
reflects morphosyntactic classes given in the official dictionary
and different solutions are described and implemented
in the well-established Vim text editor~\cite{vim}.
The tentative TP Wordnet is made available in three forms that
reflect the choices of the synsets related to each word.
In summary, this text holds potentially novel conceptualizations about,
and tools and results in analyzing, synthesizing or syntax highlighting the
TP language.
  % subj (li), obj (e), prepositions, qualification (word concatenation), 
  % sent (la), particles, mi-sina, pi, punctuation is freestyle, pre-verb,
\end{abstract}
{\bf keywords:} Constructed languages, Natural Language Processing, Syntax highlighting, Wordnet, Toki Pona

>>>>

\begin{document}
\title{Basic concepts and tools for the Toki Pona minimalist and
constructed language:
Wordnet synsets; analysis of the vocabulary; synthesis and syntax highlighting of texts}
% Anthropological computation
\author{Renato Fabbri\\
\texttt{renato.fabbri@gmail.com}\\
University of São Paulo,\\
Institute of Mathematical and Computer Sciences\\
São Carlos, SP, Brazil
}
\maketitle
\begin{abstract}
A minimalist constructed language (conlang)
is useful for experiments and comfortable for making tools.
The Toki Pona (TP) conlang is minimalist both in the vocabulary
(with only 14 letters and 124 words)
and in the $\approx10$ syntax rules.
The language is useful for being a used and somewhat established
minimalist conlang with at least hundreds of fluent speakers.
In this article, we describe current concepts and resources
for TP,
and make available Python scripted routines for
the analysis of the language,
the synthesis of texts, the specification of syntax highlighting
schemes, and the achievement of a preliminary TP
Wordnet~\cite{wordnet}.
We focus on the analysis of the basic vocabulary,
as corpus analyses were found in~\cite{corpus}.
The synthesis is based on sentence templates,
relates to context by keeping track of used words,
and renders larger texts by using a fixed number of phonemes (e.g. for poems)
and number of sentences, words and letters (e.g. for paragraphs).
Syntax highlighting 
reflects morphosyntactic classes given in the official dictionary
and different solutions are described and implemented
in the well-established Vim text editor~\cite{vim}.
The tentative TP Wordnet is made available in three forms that
reflect the choices of the synsets related to each word.
In summary, this text holds potentially novel conceptualizations about,
and tools and results in analyzing, synthesizing and syntax highlighting the
TP language.
  % subj (li), obj (e), prepositions, qualification (word concatenation), 
  % sent (la), particles, mi-sina, pi, punctuation is freestyle, pre-verb,
\end{abstract}
{\bf keywords:} Constructed languages, Natural Language Processing, Syntax highlighting, Wordnet, Toki Pona
